# Geocentric models

>This chapter introduces linear regression as a Bayesian procedure. Under a probability interpretation, which is necessary for Bayesian work, linear regression uses a Gaussian (normal) distribution to describe our golem’s uncertainty about some measurement of interest. This type of model is simple, flexible, and commonplace. Like all statistical models, it is not universally useful. But linear regression has a strong claim to being foundational, in the sense that once you learn to build and interpret linear regression models, you can more easily move on to other types of regression which are less normal.
>>p. 71

## Why normal distributions are normal
Imagine there are 1000 people standing on the center line of a soccer field and they each have a coin. They each toss their coin 16 times and with each toss, heads means one step left, and tails means one step right. After everyone has done their 16 tosses, can you imagine the distribution of where they would be on the field? It would be roughly normal centered on the center of the field.

### Normal by addition
To simulate this in R we can assign random values that are either left (-1) or right (1) for each coin toss and then add them together.
```{r, Rcode 4.1}
pos <- replicate(1000, sum(runif(16,-1,1)))

plot(density(pos))
```
Now lets visualize what each person's walk
```{r}
#load packages
library(dplyr)
library(tidyr)
library(ggplot2)
#set seed for reproducible plots
set.seed(11)
#create dataframe
pos_df <- replicate(100, runif(16, -1, 1)) %>% #simulation (turned down to 100 people)
  as_tibble() %>% # make it a tibble
  rbind(0, .) %>% # add a row of zeros above the coin tosses as a starting point
  mutate(step = 0:16) %>% #this is the step index
  gather(key, value, -step) %>% #this will convert the data into long format
  mutate(person = rep(1:100, each = 17)) %>%  #this adds a person index (17 times for the 0 and 16 coins)
  group_by(person) %>% #for each person value
  mutate(position = cumsum(value)) %>% #create a cumulative sum for each person
  ungroup()

#create plot
pos_df %>% 
  ggplot(aes(x = step, y = position, group = person)) +
  geom_vline(xintercept = c(4, 8, 16), linetype = 2) +
  geom_line(aes(color = person < 2, alpha = person < 2)) +
  scale_color_manual(values = c('skyblue4','black')) +
  scale_alpha_manual(values = c(1/5, 1)) + 
  scale_x_continuous('step number', breaks = c(0, 4, 8, 12, 16)) +
  theme(legend.position = 'none',
        panel.grid = element_blank())
```

You can see that the more steps you add, the more normal the distribution becomes. 

### Normal by multiplication

Suppose that an organism's growth rate is influenced by 12 loci each with several alleles that code for more growth and that these loci can interact with each other causing multiplicative effects
```{r, Rcode 4.2}
set.seed(11)
prod(1 + runif(12, 0, 0.1))
```
Or more explicitly:
```{r}
set.seed(11)
a = 1
b = runif(12, 0, 0.1)
c = a + b
prod(c)
```
as a density:
```{r, Rcode 4.3}
set.seed(11)
growth <- replicate(10000, prod(1 + runif(12, 0, 0.1)))
plot(density(growth))
```

Now see how the distribution behaves if you change the effect size of the loci
```{r}
set.seed(11)
big <- replicate(10000, prod(1 + runif(12, 0, 0.5)))
plot(density(big))
```
```{r}
set.seed(11)
small <- replicate(10000, prod(1 + runif(12, 0, 0.01)))
plot(density(small))
```

### Normal by log-multiplication

See how skewed the ```big``` distribution was above? If we log transform the product, what do you think we will see?
```{r, Rcode 4.5}
set.seed(11)
log_big <- replicate(10000, log(prod(1 + runif(12, 0, 0.5))))
plot(density(log_big))
```

### Using Gaussian distributions

define ontological


define epistemological


## A language for describing models

Learning the language
1. Recognize the variables you are working with. Observable variables are _data_ and unobservable variables are parameters\
2. Define the variable in terms of the other variables or as probability distributions\
3. The combination of variables and probability distributions are _joint generative models_ that can be used to simulate hypothetical observations as well as analyze real ones\

Summarize the model in a mathy way:\

\begin{equation}
    y_{i} ~ \text{Normal}(\mu_{i}, \sigma)\
    \mu_{i} ~ \beta x_{i}\
    \beta ~ \text{Normal}(0, 10)\
    \sigma ~ \text{Exponential}(1)\
    x_{i} ~ \text{Normal}(0,1)\
\end{equation}

### Re-describing the globe tossing model. 

Recall that the globe tossing model had the variables defined as:\

\begin{equation}
W ~ \text{Binomial}(N, p)\
p ~ \text{Uniform}(0, 1)
\end{equation}

These definitions have implied meanings. For example, the Binomial distribution of $W$ implies that each toss ($N$) is independent of the others. The first line of these simple models are the likelihood function from Bayes' theorem. The other lines will outline priors for variables if defined.

## Gaussian model of height

In this section we will build the beginnings of a regression model that will follow a gaussian distribution with a mean $\mu$ and standard deviation $\sigma$. The model will take in data or a predictor variable and evaluate all possible values of $\mu$ and $\sigma$ to produce a distribution of distributions basically.

### The data

Let's load up McElreath's data
```{r, Rcode 4.7}
library(rethinking)
data(Howell1)
d <- Howell1
```

Now we can see the new data frame
```{r, Rcode 4.8}
str(d)
```

```{r, Rcode 4.9}
precis(d, hist=FALSE) #rethinking package
```

And for this model we will just be using height of adults
```{r, Rcode 4.10}
d2 <- d[d$age >= 18 ,] #create a new data frame that is all rows where age is 18 or greater
```

### The Model

We want to model these heights with a Gaussian distribution. Let's see the data first
```{r, Rcode 4.11}
dens(d2$height)
```

This is pretty normal looking so we can go ahead and set the model up to use a Gaussian distirbution to describe the probability distribution of the data. Because we are still unsure about which distribution is the right one, we will leave the mean and standard deviation as variables for now.\

\begin{equation}
h_i ~ \text{Normal}(\mu, \sigma)
\end{equation}

This is our likelihood function for the model. Next we will need to define our $\mu$ and $\sigma$ variables.

\begin{equation}
\mu ~ \text{Normal}(178, 20)
\sigma ~ \text{Uniform}(0, 50)
\end{equation}

Here McElreath uses 178 cm as the mean because that is his height and spreads the 95% interval 40 cm in either direction (138 - 218cm). Let's see what this would look like

```{r, Rcode 4.12}
curve(dnorm(x, 178, 20), from = 100, to = 250)
```

Here we are looking across heights from 100cm to 250cm and as you can see the normal distribution is putting the the mean between 140 and 220. 

$\sigma$ has a uniform prior that holds its value somewhere between 0 and 50 cm. 
```{r, Rcode 4.13}
curve(dunif(x, 0, 50), from = -10, to = 60)
```

Here you can see that any value from 0 to 50 has an equal probability of being 'correct'. Now we can create something called a _Prior Predictive_ by using the priors we've defined to see how they relate to the observed heights.

```{r, Rcode 4.14}
sample_mu <- rnorm(1e4, 178, 20) #create randome draws from the mu prior
sample_sigma <- runif(1e4, 0, 50) #create random draws from the sigma prior
prior_h <- rnorm(1e4, sample_mu, sample_sigma) #use the priors of mu and sigma to infulence h
dens(prior_h)
```

### Grid approximation of the posterior distribution

Here we will approximate the posterior in detail with grid approximation for learning purposes. First will be the methods presented by McElreath and second I will add a tidyverse version adapted from [Kurz](https://bookdown.org/ajkurz/Statistical_Rethinking_recoded/linear-models.html)

```{r, Rcode 4.16}
mu_list <- seq(from = 150, to = 160, length.out = 100) #100 equally spaced values from 150 to 160
sigma_list <- seq(from = 7, to = 9, length.out = 100) #100 equally spaced values from 7 to 9
post <- expand.grid(mu = mu_list, sigma = sigma_list) #expand to a grid of all combinations of the above (100 x 100)
post$LL <- sapply(1:nrow(post), function(i) sum(
  dnorm(d2$height, post$mu[i], post$sigma[i], log = TRUE))) #calculate the log likelihood for each row of post across the data values of d2$height
post$prod <- post$LL + dnorm(post$mu, 178, 20, TRUE) + dunif(post$sigma, 0, 50, TRUE) #create a product by adding priors to log likelihood
post$prob <- exp(post$prod - max(post$prod)) #Create the posterior probability by taking each product and subtracting the maximum product (while also back transforming to cm)
```

Now we can see the posterior with:
```{r, Rcode 4.17}
contour_xyz(post$mu, post$sigma, post$prob)
```

or:
```{r, Rcode 4.18}
image_xyz(post$mu, post$sigma, post$prob)
```

In tidyverse language
```{r}
n <- 100

d_grid <- 
  as_tibble(mu <- seq(from = 150, to = 160, length.out = n),
         sigma <- seq(from = 7, to = 9, length.out = n)) 

d_grid <- d_grid %>% 
  tidyr::expand(mu, sigma)

head(d_grid)
```

```{r}
grid_function <- function(mu, sigma){
  dnorm(d2$height, mean = mu, sd = sigma, log = TRUE) %>% 
    sum()
}

d_grid <- d_grid %>% 
  mutate(log_likelihood = purrr::map2(mu, sigma, grid_function)) %>% 
  unnest() %>% 
  mutate(prior_mu = dnorm(mu, mean = 178, sd = 20, log = TRUE),
         prior_sigma = dunif(sigma, min = 0, max = 50, log = TRUE)) %>% 
  mutate(product = log_likelihood + prior_mu + prior_sigma) %>% 
  mutate(probability = exp(product - max(product)))

head(d_grid)
```

```{r}
d_grid %>% 
  ggplot(aes(x = mu, y = sigma, z = probability)) +
  geom_contour() +
  labs(x = expression(mu),
       y = expression(sigma)) +
  xlim(range(d_grid$mu)) +
  ylim(range(d_grid$sigma)) +
  theme(panel.grid = element_blank())
```

```{r}
d_grid %>% 
  ggplot(aes(x = mu, y = sigma)) +
  geom_raster(aes(fill = probability), interpolate = TRUE) +
  scale_fill_viridis_c(option = 'A') +
  labs(x = expression(mu),
       y = expression(sigma)) +
  theme(panel.grid = element_blank())
```

### Sampling from the posterior

Now that we have a posterior calculated, we can pull samples from it.
```{r, Rcode 4.19}
sample_rows <- sample(1:nrow(post), size = 1e4, replace = TRUE, prob = post$prob) #sample rows by their probability
sample_mu <- post$mu[sample_rows] #index samples of mu by the rows sampled
sample_sigma <- post$sigma[sample_rows] #index samples of sigma by rows sampled
```

Now we can plot these samples to see where the most common values for $\mu$ and $\sigma$ might overlap
```{r, Rcode 4.20}
plot(sample_mu, sample_sigma, cex = 1, pch = 16, col = col.alpha(rangi2, 0.1))
```

We can also look at the samples individually
```{r, Rcode 4.21}
dens(sample_mu)
dens(sample_sigma)
```

And we can get summary points from these sample distributions just like before
```{r, Rcode 4.22}
PI(sample_mu)
PI(sample_sigma)
```

(tidyverse)
```{r}
d_grid_samples <- d_grid %>% 
  sample_n(size = 1e4, replace = TRUE, weight = probability)

d_grid_samples %>% 
  ggplot(aes(x = mu, y = sigma)) +
  geom_point(size = 0.9, alpha = 1/15) +
  scale_fill_viridis_c() +
  labs(x = expression(mu[samples]),
       y = expression(sigma[samples])) +
  theme(panel.grid = element_blank())
  
```

```{r}
d_grid_samples %>% 
  select(mu, sigma) %>% 
  gather() %>% 
  ggplot(aes(x = value)) +
  geom_density(fill = 'grey35', size = 0) +
  scale_y_continuous(NULL, breaks = NULL) +
  xlab(NULL) +
  theme(panel.grid = element_blank()) +
  facet_wrap(~key, scales = 'free')
```

```{r}
library(tidybayes)

d_grid_samples %>% 
  select(mu, sigma) %>% 
  gather() %>% 
  group_by(key) %>% 
  mode_hdi(value)
```

```{r}
d_grid_samples %>% 
  select(mu, sigma) %>% 
  gather() %>% 
  group_by(key) %>% 
  median_qi(value, .width = 0.5)
```

### Finding the posterior distribution with ```quap```

Recall that _quadratic approximation_ uses the _maximum a posteriori_ (MAP) as the peak of the distribution and creates the shape of the distribution as a negative quadratic based on the slope near the peak.

To set up, lets rerun the data input
```{r, Rcode 4.26}
data(Howell1)
d <- Howell1
d2 <- d[d$age >= 18,]
```

Now we have to define our model parameters in R code
```{r, Rcode 4.27}
flist <- alist(
height ~ dnorm(mu, sigma),
mu ~ dnorm(178, 20),
sigma ~ dunif(0, 50)
)
```

And then we can fit the model to the data
```{r, Rcode 4.28}
m4.1 <- quap(flist, data = d2)
```

and the posterior is available to view here
```{r, Rcode 4.29}
precis(m4.1)
```

Changing the prior for $\mu$ to be quite narrow
```{r, Rcode 4.31}
m4.2 <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu ~ dnorm(178, 0.1),
    sigma ~ dunif(0, 50)
  ), data = d2)

precis(m4.2)
```

notice here that because the prior was so narrow for $\mu$, the mean did not change much from the prior value. $\sigma$ on the other hand has vastly changed from the previous example even though its prior remained the same.

### Sampling with ```quap```

Because quadratic approximation is inherently multidimensional, we need to access to the different covariances between pairs of parameters. Here is the variance-covariance matrix for the first model.
```{r, Rcode 4.32}
vcov(m4.1)
```

The above contains both the variance of each parameter but also the correlated change expected in the other parameters. To see them separately:
```{r, Rcode 4.33}
diag(vcov(m4.1)) #variance of each parameter (sqrt = sd)

cov2cor(vcov(m4.1)) #Correlation matrix of the parameters 
```

The correlation matrix above indicates that $\mu$ and $\sigma$ are very far from correlated. Now let's pull vectors of values from a multi-dimenstional distribution (made easy with ```rethinking```)
```{r, Rcode 4.34}
post <- extract.samples(m4.1, n = 1e4)

head(post)
```

Now we can compare to the posterior samples to the MAP values
```{r, Rcode 4.35}
precis(post)
```

## Linear Prediction

Now that we have a model for height, we can get to the 'regression' part where we add a _predictor variable_. Let's add weight to our height model and see what that looks like

```{r, Rcode 4.37}
data(Howell1); d <- Howell1; d2 <- d[d$age >= 18, ]
plot(d2$height ~ d2$weight)
```

Seems to be a pretty clear relationship between height and weight. The linear model strategy will take the predictor variable (weight) and assume that it has a constant additive relationship to the mean of the outcome (height). Now we have to put our predictor into the model structure.\


\begin{equation}
    h_{i} ~ \text{Normal}(\mu_{i}, \sigma)
    \mu_{i} = \beta (x_{i} - \overline{x})
    \alpha ~ \text{Normal}(178, 20)
    \beta ~ \text{Uniform}(0, 10)
    \sigma ~ \text{Uniform}(0, 50)
\end{equation}

#### Probability of the data (first line of the model)

So now that both height $h_{i}$ and the mean $\mu_{i}$ are indexed by $i$ it means that each data point or row in the dataframe influences the value of $\mu$ and therefore $h$.

#### Linear model (second line of the model)

The value of $\mu_{i}$ is now made from a combination of $\alpha$, $\beta$, and observed weight data ($x$) where $\overline{x}$ describes the mean of all weight values observed.

There are two parts to this line in the model that make it 'linear'\
1. When $x_{i}$ = $\overline{x}$ then the $\beta$ value is 0 which makes $\alpha$ an 'intercept'.\
2. When $x_{i}$ increases by one unit, the expected change in height ($h_{i}$) is explained by $\beta$ or the 'slope'.\

#### Priors (lines 3-5 of the model)

The $\alpha$ and $\sigma$ priors is basically the same as before (except now the $\mu$ prior is now $\alpha$). The $\beta$ prior is set to have a mean of 0 on purpose because we don't know for sure what the relationship is between weight and height (even though the plot suggests a positive slope value). To illustrate this point, McElreath simulates the prior predictive distribution using observed weights.

```{r, Rcode 4.38}
set.seed(2971)
N <- 100 # 100 lines
a <- rnorm(N, 178, 20) # alpha prior
b <- rnorm(N, 0, 10) # beta prior
```

Let's see how this random selection from the priors looks
```{r, Rcode 4.39}
plot(NULL, xlim = range(d2$weight), ylim = c(-100, 400),
     xlab = 'weight', ylab = 'height')
abline(h = 0, lty = 2) #dashed line at height = 0
abline(h = 272, lty = 1, lwd = 0.5) #solid line at world's tallest person
mtext('b ~ dnorm(0, 10)') #plot title
xbar <- mean(d2$weight) #average weight
for(i in 1:N){
  curve(a[i] + b[i]*(x - xbar), #linear equation for each pair
  from = min(d2$weight), to = max(d2$weight), add = TRUE,
  col = col.alpha('black',0.2))
}
```

This is clearly absurd as no adult should be near 0 height and weigh 30 kg. We can fix this by holding $\beta$ positive with a log-normal distribution

```{r, Rcode 4.40}
b <- rlnorm(1e4, 0, 1)
dens(b, xlim = c(0,5), adj = 0.1)
```

Now we can use this new $\beta$ prior in the last plot
```{r, Rcode 4.41}
set.seed(2971)
N <- 100 # 100 lines
a <- rnorm(N, 178, 20) # alpha prior
b <- rlnorm(N, 0, 1) # NEW beta prior

plot(NULL, xlim = range(d2$weight), ylim = c(-100, 400),
     xlab = 'weight', ylab = 'height')
abline(h = 0, lty = 2) #dashed line at height = 0
abline(h = 272, lty = 1, lwd = 0.5) #solid line at world's tallest person
mtext('log(b) ~ dnorm(0, 1)') #plot title
xbar <- mean(d2$weight) #average weight
for(i in 1:N){
  curve(a[i] + b[i]*(x - xbar), #linear equation for each pair
  from = min(d2$weight), to = max(d2$weight), add = TRUE,
  col = col.alpha('black',0.2))
}
```

This new plot seems far more plausible for an adult population.

### Finding the posterior distribtuion

Just like before we can use the ```quap``` function to get a quadratic approximation or alternatively, we can use brms to fit the model.

```quap``` method

```{r, Rcode 4.42}
#re load the data
data(Howell1) ; d <- Howell1; d2 <- d[d$age >= 18,]

#define the mean weight of observed values
xbar <- mean(d2$weight)

#fit the model
m4.3 <- quap(
  alist(
    height ~ dnorm(mu, sigma), #height likelihood distribution
    mu <- a + b*(weight - xbar), #linear equation
    a ~ dnorm(178, 20), #alpha prior
    b ~ dlnorm(0, 1), #beta prior
    sigma ~ dunif(0, 50) #sigma prior
  ), data = d2
)

precis(m4.3)
```

brms method

```{r}
library(brms)
b4.3_priors <- get_prior(data = d2, formula = height ~ 1 + weight)

b4.3 <- brm(data = d2, #data to be used 
            family = gaussian, #likelihood of height
            height ~  1 + weight, #linear equation
            prior = b4.3_priors,
            iter = 41000, warmup = 40000, chains = 4, cores = 4, seed = 11) #model mechanics
```
The plot

```{r}
plot(b4.3)
```

The alpha values (b_intercept in brms) appear to be quite different here. This is because we haven't centered the weight values yet ($x - \overline{x}$), so we can do that and run it again.

```{r}
d2 <- d2 %>% 
  mutate(weight_c = weight - mean(weight))

b4.3b_priors <- get_prior(data = d2, formula = height ~ 1 + weight_c)

b4.3b <- brm(data = d2, #data to be used 
            family = gaussian, #likelihood of height
            height ~  1 + weight_c, #linear equation
            prior = b4.3b_priors,
            iter = 41000, warmup = 40000, chains = 4, cores = 4, seed = 11) #model mechanics

plot(b4.3b)
```


### Interpreting the posterior distribution

>[The model] can only report the posterior distribution. This is the right answer to the question you asked. But it's your responsibility to process the answer and make sense of it.
>
>---McElreath, p.98

You can make sense of model outputs in two ways:\
1. read summary tables\
2. plot simulations\

Reading summary tables becomes rather complicated as the models become more complex.

#### Tables of marginal distributions

```{r, Rcode 4.44}
precis(m4.3)
```

In this table, the first row is the quadratic approximation for $\alpha$, second is $\beta$, and finally $\sigma$. 

repeated for our brms model
```{r}
round(posterior_summary(b4.3b, probs = c(0.055,0.945))[1:3,], 2)
```


Now let's see the variance-covariance matrix
```{r, Rcode 4.45}
round(vcov(m4.3), 3)
```

repeated for brms model
```{r}
posterior_samples(b4.3b) %>% 
  select(-lp__) %>% 
  cor() %>% 
  round(digits = 2)
```

Visual approach
```{r}
pairs(m4.3)
```

*very little covariance here

#### Plotting the posterior inference against the data

```{r, Rcode 4.46}
plot(height ~ weight, data = d2, col = rangi2) #plot the data points
post <- extract.samples(m4.3) #extract samples from the posterior
a_map <- mean(post$a) #get the mean of the samples of alpha
b_map <- mean(post$b) #get the mean of the samples of beta
curve(a_map + b_map*(x - xbar), add = TRUE)
```

This line is very plausible but ther are others that are equally plausible

```{r}
d2 %>% 
  ggplot(aes(x = weight, y = height)) +
  geom_abline(intercept = fixef(b4.3)[1], 
              slope = fixef(b4.3)[2]) +
  geom_point(shape = 1, size = 2, color = 'royalblue') +
  theme_bw() + 
  theme(panel.grid = element_blank())
```


#### Adding uncertainty around the mean

There are many combinations of $\alpha$ and $\beta$ that could be plausible
```{r, Rcode 4.47}
post <- extract.samples(m4.3)
post[1:5,]
```

We will also alter the amount of observations to see how it influences the spread of these combinations.

```{r, Rcode 4.48}
N <- 10 #lets pull 10 data points
dN <- d2[1:N, ]
mN <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b*(weight - mean(weight)),
    a ~ dnorm(178, 20),
    b ~ dlnorm(0, 1),
    sigma ~ dunif(0, 50)
  ), data = d2)


```

```{r, Rcode 4.49}
#and plot 20 samples
post <- extract.samples(mN, n=20)

#display data and sample size
plot(dN$weight, dN$height, xlim = range(d2$weight), ylim = range(d2$height),
     col = rangi2, xlab = 'weight', ylab = 'height',)
mtext(concat('N = ',N))

#add sample lines
for(i in 1:20){
  curve(post$a[i] + post$b[i]*(x - mean(dN$weight)),
        col=col.alpha('black',0.3), add = TRUE)
}
```

You can see that the extreme weight values are far more spread out. This pattern disappears when the model has more data to work with and becomes more confident about the means.
```{r}
N <- 300 #lets pull 300 data points
dN <- d2[1:N, ]
mN <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b*(weight - mean(weight)),
    a ~ dnorm(178, 20),
    b ~ dlnorm(0, 1),
    sigma ~ dunif(0, 50)
  ), data = d2)

#and plot 20 samples
post <- extract.samples(mN, n=25)

#display data and sample size
plot(dN$weight, dN$height, xlim = range(d2$weight), ylim = range(d2$height),
     col = rangi2, xlab = 'weight', ylab = 'height',)
mtext(concat('N = ',N))

#add sample lines
for(i in 1:25){
  curve(post$a[i] + post$b[i]*(x - mean(dN$weight)),
        col=col.alpha('black',0.3), add = TRUE)
}
```

#### Plotting regression intervals and contours

Let's use 50 kg as a focal weight and grab some samples
```{r, Rcode 4.50}
post <- extract.samples(m4.3)
mu_at_50 <- post$a + post$b * (50 - xbar)
```

Now we can look at the distribution of plausible heights for a person of 50kg weight
```{r, Rcode 4.51}
dens(mu_at_50, col = rangi2, xlab = 'mu|weight = 50kg')
```

The height of someone that weighs 50 kg is almost certainly between 158 and 160.5 cm.

```{r, Rcode 4.52}
PI(mu_at_50, prob = 0.89)
```

and 89% of the ways the model can produce the data suggests that the average height of a 50kg person is between 159 and 160cm.

Now what can we do about the average slope ($\beta$) from the model?
```{r, Rcode 4.53}
mu <- link(m4.3)
str(mu)
```

Now we have a large matrix that has 1000 samples (rows) for 352 data points (observations). But we want to have increments of 1 kg steps across the x axis.

```{r, Rcode 4.54}
weight.seq <- seq(from = 25, to = 70, by =1)

mu <- link(m4.3, data = data.frame(weight = weight.seq))
str(mu)
```

Now our new matrix has 1000 samples across 46 weight steps. Let's visualize the uncertainty of each weight value.
```{r, Rcode 4.55}
plot(height ~ weight, d2, type = 'n') #type 'n' hides the data
for (i in 1:100){
  points(weight.seq, mu[i,], pch = 16, col = col.alpha(rangi2, 0.1))
}
```

Now we can compute a mean $\mu$ value and the distribution for each weight value
```{r, Rcode 4.56}
mu_mean <- apply(mu, 2, mean)
mu_PI <- apply(mu, 2, PI, prob = 0.89)
```

And add them to the plot
```{r, Rcode 4.57}
plot(height ~ weight, data = d2, col = col.alpha(rangi2, 0.5))
shade(mu_PI, weight.seq, col = 'grey80')
lines(weight.seq, mu_mean)
```

#### Prediction Intervals

Now that we have a good handle on $\mu_{i}$ we can combine it with $\sigma$ to create estimates of height rather than just the mean. 

```{r, Rcode 4.59}
sim.height <- sim(m4.3, data = list(weight = weight.seq))
str(sim.height)
```

```{r, Rcode 4.60}
height.PI <- apply(sim.height, 2, PI, prob = 0.89)
```

```{r, Rcode 4.61}
plot(height ~ weight, d2, type = 'n')
shade(height.PI, weight.seq, col = 'grey85')
shade(mu_PI, weight.seq, col = 'grey60')
points(x = d2$weight, y = d2$height, pch = 16, col = col.alpha(rangi2, 0.5))

lines(weight.seq, mu_mean)


```

*Note: shade() seeems to be broken for me in that without specifying a colour, it is completely transparent*

If you want to smooth the edges of the interval, you can add more samples to the ```sim.height``` controls. 
```{r, Rcode 4.62}
sim.height <- sim(m4.3, data=list(weight = weight.seq), n=1e4)
height.PI <- apply(sim.height, 2, PI, prob = 0.89)

plot(height ~ weight, d2, type = 'n')
#plot(height ~ weight, data = d2, col = col.alpha(rangi2, 0.5))
shade(height.PI, weight.seq, col = 'grey85')
shade(mu_PI, weight.seq, col = 'grey60')
points(x = d2$weight, y = d2$height, pch = 16, col = col.alpha(rangi2, 0.5))
lines(weight.seq, mu_mean)
```

## Curves from lines

There are two common methods to build curves from lines. One is _Polynomial Regression_ and the other is _B-Splines_.

### Polynomial regression

This method uses powers of a variable (squares & cubes) as extra pedictors. Let's grab all the height and weight data, not just adults.
```{r, Rcode 4.64}
data(Howell1)
d <- Howell1

plot(height~weight, d)
```

This is quite the curve. We can try to model it with a parabolic model:\

\begin{equation}
\mu_{i} = \alpha + \beta_{1} x_{i} + \beta_{2} x_{i}^{2}
\end{equation}

where now each weight observation $x_{i}$ is used twice. Once as before as a linear fit and a second time squared to produce the curvature of the line defined as $\beta_{2}$

To fit the model to the data we will need to standardize the predictor variable.
```{r, Rcode 4.65}
d$weight_s <- (d$weight - mean(d$weight))/sd(d$weight)
d$weight_s2 <- d$weight_s^2
m4.5 <- quap(
  alist(
    height ~ dnorm(mu, sigma), #same as before
    mu <- a + b1*weight_s + b2*weight_s2, #adding the b2 and squared term here
    a ~ dnorm(178, 20), #same
    b1 ~ dlnorm(0, 1), #same as beta prior before
    b2 ~ dnorm(0, 1), #new prior that could be positive or negative
    sigma ~ dunif(0, 50)
  ), data = d
)
```

Now try and interpret the results of the summary table
```{r, Rcode 4.66}
precis(m4.5)
```

Not very straight forward how the numbers in the table relate to the data

Let's find the mean for $\mu$ again now with $\beta_{1}$ and $\beta_{2}$ in the equation.

```{r, Rcode 4.67}
weight.seq <- seq(from = -2.2, to = 2, length.out = 30) # grab 30 equally spaced weight values
pred_dat <- list(weight_s = weight.seq, weight_s2 = weight.seq^2) #new weight values
mu <- link(m4.5, data = pred_dat) #get new estimates from new data
mu.mean <- apply(mu, 2, mean) #calculate the mu mean for the line
mu.PI <- apply(mu , 2, PI, prob = 0.89) #calculate a 89% interval around the mean for mu
sim.height <- sim(m4.5, data = pred_dat) #calculate estimates for heights based on mu and sigma with new weight values
height.PI <- apply(sim.height, 2, PI, prob = 0.89) #calculate 89% intervals of height
```

```{r, Rcode 4.68}
plot(height~weight_s, d, type = 'n')
lines(weight.seq, mu.mean)
shade(mu.PI, weight.seq)
shade(height.PI, weight.seq)
points(x = d$weight_s, y = d$height, col = col.alpha(rangi2, 0.5))
```

Now let's try adding a cubic predictor term $\beta_{3}x_{i}^3$

```{r, Rcode 4.69}
d$weight_s3 <- d$weight_s^3
m4.6 <- quap(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b1*weight_s + b2*weight_s2 + b3*weight_s3,
    a ~ dnorm(178, 20),
    b1 ~ dlnorm(0, 1),
    b2 ~ dnorm(0, 1),
    b3 ~ dnorm(0,1),
    sigma ~ dunif(0, 50)
  ), data = d
)
```


```{r}
weight.seq <- seq(from = -2.2, to = 2, length.out = 30) # grab 30 equally spaced weight values
pred_dat <- list(weight_s = weight.seq, weight_s2 = weight.seq^2, weight_s3 = weight.seq^3) #new weight values
mu <- link(m4.6, data = pred_dat) #get new estimates from new data
mu.mean <- apply(mu, 2, mean) #calculate the mu mean for the line
mu.PI <- apply(mu , 2, PI, prob = 0.89) #calculate a 89% interval around the mean for mu
sim.height <- sim(m4.6, data = pred_dat) #calculate estimates for heights based on mu and sigma with new weight values
height.PI <- apply(sim.height, 2, PI, prob = 0.89) #calculate 89% intervals of height
```

```{r}
plot(height~weight_s, d, type = 'n')
lines(weight.seq, mu.mean)
shade(mu.PI, weight.seq)
shade(height.PI, weight.seq)
points(x = d$weight_s, y = d$height, col = col.alpha(rangi2, 0.5))
```

And here is how you would plot the above with the original x-axis
```{r, Rcode 4.70}
plot(height ~ weight_s, d, col=col.alpha(rangi2, 0.5),  type = 'n', xaxt = 'n') #turn off x axis text
#Rcode 4.71
at <- c(-2, -1, 0, 1, 2)
labels <- at*sd(d$weight) + mean(d$weight)
axis(side = 1, at = at, labels = round(labels, 1))
lines(weight.seq, mu.mean)
shade(mu.PI, weight.seq)
shade(height.PI, weight.seq)
points(x = d$weight_s, y = d$height, col = col.alpha(rangi2, 0.5))
```

### Splines

Splines are smooth functions that are built from smaller component functions that usually take on a wiggly appearance.

```{r, Rcode 4.72}
#load cherry blossom bloom time over a thousand years
library(rethinking)
data(cherry_blossoms)
d <- cherry_blossoms
precis(d)
```
Let's see the bloom time over the years
```{r}
plot(x=d$year, y=d$doy)
abline(h=mean(d$doy, na.rm = T), col = 'red', lty = 2)
```

Similar to the polynomial method, B-splines will break the predictor (year) into chunks with different basis functions to fit the data in that chunk. When we stitch the chunks together, we often get a wiggly line through the data.

The new linear function for $\mu_{i}$ will look a bit different.\

\begin{equation}
\mu_{i} = \alpha + w_{1}\beta_{i,1} + w_{2}\beta_{i,2} + w_{3}\beta_{i,3} + ...
\end{equation}

Here, the $\beta$ values are for each basis function and the $w$ parameters are the weights for each function.

For the cherry blossom data we will split the data up into 15 quantiles and have the basis functions operate within each quantile.

```{r, Rcode 4.73}
d2 <- d[complete.cases(d$doy), ] #remove any NA values
num_knots <- 15 #number of sections in the data
knot_list <- quantile(d2$year, probs = seq(0, 1, length.out = num_knots))

knot_list
```

So now that we have the break points in the year range of the data we can assign a polynomial degree. This degree determines how many basis functions will be interacting in each year range. If we choose a cubic spline, four basis functions will be active in each year range.

```{r, Rcode 4.74}
library(splines)
B <- bs(d2$year, knots = knot_list[-c(1,num_knots)], degree = 3, intercept = TRUE)
```

We just created a matrix that contains 827 rows and 17 columns. Each row is a year in ```d2``` and each column is a basis function. Let's see what it looks like (fig 4.13a)

```{r, Rcode 4.75}
plot(NULL, xlim = range(d2$year), ylim=c(0,1), xlab='year', ylab='basis')
for(i in 1:ncol(B)){
  lines(d2$year, B[,i])
}
points(x = knot_list, y = rep(0.9, length(knot_list)), col = 'red')
```

Now we need to calculate the parameter weights ($w$) for each basis function. To do this we need to run the model.

\begin{equation}
D_{i} \sim \text{Normal}(\mu_{i},\sigma)
\mu_{i} = \alpha + \sum_{k=1}^{K}w_{k}B_{k,i}
\alpha \sim \text{Normal}(100,10)
w_{j} \sim \text{Normal}(0,10)
\sigma \sim \text{Exponential}(1)
\end{equation}

```{r, Rcode 4.76}
m4.7 <- quap(
  alist(
    D ~ dnorm(mu, sigma),
    mu <- a + B %*% w,
    a ~ dnorm(100, 10),
    w ~ dnorm(0, 10),
    sigma ~ dexp(1)
  ), data = list(D = d2$doy, B=B),
  start = list(w=rep(0, ncol(B)))
)
```

Now we can extract some samples and look at the basis weights (fig 4.13b)

```{r, Rcode 4.77}
post <- extract.samples(m4.7)
w <- apply(post$w, 2, mean)
plot(NULL, xlim=range(d2$year), ylim=c(-6,6), xlab = 'year', ylab='basis * weight')
for(i in 1:ncol(B)){
  lines(d2$year, w[i]*B[,i])
}
points(x = knot_list, y = rep(4.9, length(knot_list)), col = 'red')

```

Now we can calculate a posterior interval for $\mu$ at each year (fig 4.13c)
```{r, Rcode 4.78}
mu <- link(m4.7)
mu_PI <- apply(mu, 2, PI, prob = 0.97)
mu_mean <- apply(mu, 2, mean)
plot(d2$year, d2$doy, col = col.alpha(rangi2, 0.3), pch = 16)
shade(mu_PI, d2$year, col = col.alpha('black',0.5))
lines(d2$year, mu_mean)
```

A whole other class of similar models are called _GAMs_ (Generalized Additive Models) and are worth looking into if you have time.