# Conditional manatees

## Building an interaction
$R$ = ruggedness, $G$ = GDP, $C$ = continent, $U$ = unobserved variables.
```{r}
library(dagitty)
dag_8.1 <- dagitty("dag{
                   R -> G
                   C -> G
                   U -> G
                   U -> R
                   }")
coordinates(dag_8.1) <- list(y = c(R = 0, G = 0, C = 0, U = 1),
                             x = c(R = 0, G = 1, U = 1, C = 2))
drawdag(dag_8.1)
```

$G = f(R,C)$ 

### Making a rugged model

```{r, Rcode 8.1}
library(rethinking)
data(rugged)
d <- rugged

#log transform GDP
d$log_gdp <- log(d$rgdppc_2000)

#only include countries with GDP data
dd <- d[complete.cases(d$rgdppc_2000),]

#rescale variables
dd$log_gdp_std <- (dd$log_gdp) / mean(dd$log_gdp) # values of 1 is average
dd$rugged_std <- (dd$rugged) / max(dd$rugged) # values range from 0 to max ruggedness (1)

```

Basic model
$$\text{log}(y_{i}) \sim \text{Normal}(\mu_{i}, \sigma)\\
\mu_{i} = \alpha + \beta(r_{i} - \overline{r})\\
\alpha \sim \text{Normal}|(1,1)\\
\beta \sim \text{Normal}(0, 1)\\
\sigma \sim \text{Exponential}(1)$$

In R:
```{r, Rcode 8.2}
m8.1 <- quap(
  alist(
    log_gdp_std ~ dnorm(mu, sigma),
    mu <- a + b*(rugged_std - 0.215),
    a ~ dnorm(1, 1),
    b ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ), data = dd
)
```

Sample priors:
```{r, Rcode 8.3}
set.seed(11)
prior <- extract.prior(m8.1)

#set plot
plot(NULL, xlim=c(0,1), ylim=c(0.5, 1.5), xlab = "ruggedness", ylab = "log GDP")
abline(h=min(dd$log_gdp_std), lty = 2)
abline(h=max(dd$log_gdp_std), lty = 2)

#draw lines from prior
rugged_seq <- seq(from = -0.1, to = 1.1, length.out=30)
mu <- link(m8.1, post = prior, data = data.frame(rugged_std=rugged_seq))
for(i in 1:50){
  lines(rugged_seq, mu[i,], col=col.alpha('black',0.3))
  }
```

$\alpha$ is too wild. intercept should be somewhere around where the mean of ruggedness hits 1 on the log GDP scale so adjust to Normal(1, 0.1).

$\beta$ is also out of control. we need something (positive or negative) that spans the difference between the dashed lines

Slope should be $\pm 0.6$ which is the differece between the maximum and minimum values of GDP
```{r}
max(dd$log_gdp_std) - min(dd$log_gdp_std)
```

```{r, echo=FALSE}
#set plot
plot(NULL, xlim=c(0,1), ylim=c(0.5, 1.5), xlab = "ruggedness", ylab = "log GDP")
abline(h=min(dd$log_gdp_std), lty = 2)
abline(h=max(dd$log_gdp_std), lty = 2)

#draw lines from prior
rugged_seq <- seq(from = -0.1, to = 1.1, length.out=30)
mu <- link(m8.1, post = prior, data = data.frame(rugged_std=rugged_seq))
for(i in 1:50){
  lines(rugged_seq, mu[i,], col=col.alpha('black',0.3))
  }

abline(a=1.3, b = -0.6, lwd = 3, col = rangi2)
abline(a = 0.7, 0.6, lwd = 3, col = rangi2)

```

```{r}
#proportion of slopes greater than 0.6
sum(abs(prior$b) > 0.6) / length(prior$b)
```

Let's fix the model
```{r}
m8.1 <- quap(
  alist(
    log_gdp_std ~dnorm(mu,sigma),
    mu <-a+b*(rugged_std-0.215),
    a ~dnorm(1,0.1),
    b ~dnorm(0,0.3),
    sigma ~dexp(1)
  ), data = dd
)

```

```{r, echo=FALSE}
set.seed(11)
prior <- extract.prior(m8.1)


#set plot
plot(NULL, xlim=c(0,1), ylim=c(0.5, 1.5), xlab = "ruggedness", ylab = "log GDP")
abline(h=min(dd$log_gdp_std), lty = 2)
abline(h=max(dd$log_gdp_std), lty = 2)

#draw lines from prior
rugged_seq <- seq(from = -0.1, to = 1.1, length.out=30)
mu <- link(m8.1, post = prior, data = data.frame(rugged_std=rugged_seq))
for(i in 1:50){
  lines(rugged_seq, mu[i,], col=col.alpha('black',0.3))
  }

abline(a=1.3, b = -0.6, lwd = 3, col = rangi2)
abline(a = 0.7, 0.6, lwd = 3, col = rangi2)

```

```{r}
precis(m8.1)
```


No association seen yet

### Adding an indicator isn't enough

Update $\mu$ 

$$\mu_{i} = \alpha_{CID[i]} + \beta(r_{i} - \overline{r})$$



```{r}
#make an index variable for Africa (1) and other continents (2)
dd$cid <- ifelse(dd$cont_africa == 1, 1, 2)
```

Now update the model
```{r}
m8.2 <- quap(
  alist(
    log_gdp_std ~ dnorm(mu, sigma),
    mu <-a[cid] + b * (rugged_std - 0.215),
    a[cid] ~ dnorm(1, 0.1),
    b ~ dnorm(0, 0.3),
    sigma ~dexp(1)
  ), data = dd
)
```


```{r}
compare(m8.1, m8.2)
```

```{r}
precis(m8.2, depth = 2)
```

```{r}
post <- extract.samples(m8.2)
diff_a1_a2 <- post$a[,1] - post$a[,2]
PI(diff_a1_a2)
```

```{r}
rugged.seq <- seq(from = -0.1, to = 1.1, length.out = 30)
mu.NotAfrica <- link(m8.2, data = data.frame(cid=2, rugged_std=rugged.seq))
mu.Africa <- link(m8.2, data = data.frame(cid = 1, rugged_std = rugged.seq))
mu.NotAfrica_mu <- apply(mu.NotAfrica, 2, mean)
mu.NotAfrica_ci <- apply(mu.NotAfrica, 2, PI, prob = 0.97)
mu.Africa_mu <- apply(mu.Africa, 2, mean)
mu.Africa_ci <- apply(mu.Africa, 2, PI)

plot(NULL, xlim=c(0,1), ylim=c(0.5, 1.5), xlab = "ruggedness", ylab = "log GDP")
points(dd$rugged_std, dd$log_gdp_std, col = dd$cid, pch = 16)
lines(rugged.seq, mu.Africa_mu, lwd = 2, col = 1)
shade(mu.Africa_ci, rugged.seq)
lines(rugged.seq, mu.NotAfrica_mu, lwd = 2, col = 2)
shade(mu.NotAfrica_ci, rugged.seq, col = col.alpha(2, 0.3))

```

### Adding an interaction does work

$$\mu_{i} = \alpha_{CID[i]} + \beta_{CID[i]}(r_{i} - \overline{r})$$

```{r}
m8.3 <- quap(
  alist(
     log_gdp_std ~ dnorm(mu, sigma),
    mu <-a[cid] + b[cid] * (rugged_std - 0.215),
    a[cid] ~ dnorm(1, 0.1),
    b[cid] ~ dnorm(0, 0.3),
    sigma ~dexp(1)
  ), data = dd
)

precis(m8.3, depth = 2)
```

```{r}
compare(m8.1, m8.2, m8.3, func=PSIS)
```

```{r}
plot(PSIS(m8.3, pointwise = TRUE)$k)
```

### Plotting the interaction

```{r}
par(mfrow=c(1,2))
# plot Africa - cid = 1
d.A1 <-dd[dd$cid == 1,]
plot(d.A1$rugged_std, d.A1$log_gdp_std, pch=16, col=rangi2,
xlab="ruggedness (standardized)",ylab="log GDP (as proportion of mean)",
xlim=c(0,1) )
mu <-link(m8.3,data=data.frame(cid=1,rugged_std=rugged_seq))
mu_mean <-apply(mu,2,mean)
mu_ci <-apply(mu,2,PI,prob=0.97)
lines( rugged_seq,mu_mean,lwd=2)
shade( mu_ci,rugged_seq,col=col.alpha(rangi2,0.3))
mtext("African nations")
# plotnon-Africa-cid=2
d.A0 <-dd[dd$cid==2,]
plot( d.A0$rugged_std,d.A0$log_gdp_std,pch=1,col="black",
xlab="ruggedness (standardized)",ylab="logGDP(asproportionofmean)",
xlim=c(0,1) )
mu <-link(m8.3,data=data.frame(cid=2,rugged_std=rugged_seq))
mu_mean <-apply(mu,2,mean)
mu_ci <-apply(mu,2,PI,prob=0.97)
lines( rugged_seq,mu_mean,lwd=2)
shade( mu_ci,rugged_seq)
mtext("Non-African nations")
```

## Symmetry of interactions
